---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Add your name here"
date: "Add the date here"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```
## Exercises

```{r}
library(magrittr)     # the pipe
library(tidyverse)    # for data wrangling + visualization
library(tidymodels)   # for modeling
library(gt)           # for making display tables
library(gtExtras)     # helper functions for beautiful tables
library(DataExplorer) #
```

### Exercise 1

```{r}
view(the_tate)
view(the_tate_artists)
```

```{r}
the_tate %>% dplyr::summarize(
  no_artists = n_distinct(artistId),
  max_period_year = max(year, na.rm = TRUE),
  min_period_year = min(year, na.rm = TRUE),
  max_period_acquisitionYear = max(acquisitionYear, na.rm = TRUE),
  min_period_acquisitionYear = min(acquisitionYear, na.rm = TRUE),
  
)
```

```{r}
the_tate %>% DataExplorer::introduce() 
the_tate %>% DataExplorer::plot_missing()
```

The `the_tate` dataset has 3342 unique artists who worked from 1545 to 2012. The works were acquired between the years 1823 and 2013 .

### Exercise 2

```{r}
ct <-  the_tate %>%   dplyr::select(artistId,title) %>% dplyr::filter(title %in% c("[no title]","Untitled","Blank")) 
table(ct)
table(the_tate$id,the_tate$year,useNA = "ifany")

#— Number of records with missing dates —
ct3 <-  sum(is.na(the_tate$year)) 
ct4 <- the_tate %>% dplyr::select(artistId , year) %>% dplyr::filter(is.na(year))

artists_with_missing_years <- ct4 %>% dplyr::group_by(artistId) %>% dplyr::summarize(
  missing_no = n()
) %>% dplyr::arrange(desc(missing_no))

artists_with_missing_years <- artists_with_missing_years %>% dplyr::mutate(pct_missing_years = (missing_no/sum(missing_no))*100,cumsum(pct_missing_years))

count(distinct(ct4,artistId))
count(ct4)
```

How number of works with missing dates is 5397.

The number of artists whose works have missing dates is 462.

It would require resolving missing year data for only 11 artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as MAR.

### Exercise 3

```{r}
artist_works <- the_tate %>% dplyr::group_by(artistId) %>% dplyr::summarize(
  number_of_works = n()
) %>% dplyr::arrange(desc(number_of_works))

artist_works <- artist_works %>% dplyr::mutate(pct_of_work = (number_of_works/sum(number_of_works))*100)

artist_works %>% gt() %>% 
  gt_theme_538() 

the_tate %>% dplyr::filter(artistId %in% c(558,2121)) %>% dplyr::select(artistId,artist) %>% distinct(artistId,artist) 
```

The artist with the most works in the Tate collection is Turner, Joseph Mallord William with artistId 558.

The artist with the tenth-most works in the Tate collection is Warhol, Andy with artistId = 2121.

### Exercise 4

The artist with the greatest number of works in the Tate collection represent 56.919697692% of the total number of works

### Exercise 5

```{r}
total_artist_title_pairs <- the_tate %>% dplyr::select(artistId , title) %>% tibble::as_tibble() %>% count()
total_artist_title_pairs_distinct <- the_tate %>% dplyr::select(artistId , title) %>% dplyr::distinct() %>% count()
#No. of Total Rows
print(total_artist_title_pairs)
#No. of Distinct Rows
print(total_artist_title_pairs_distinct)
#No. of Duplicate Rows
print(total_artist_title_pairs - total_artist_title_pairs_distinct)
```

There are 23705 duplicate artist-title pairs

### Exercise 6

```{r}
the_tate_area <- the_tate %>% dplyr::select(artist,artistId,width,height) %>% dplyr::mutate(area = (width*height)/100) %>% tidyr::drop_na(area) %>% dplyr::arrange(desc(area))

dplyr::slice_head(the_tate_area)
dplyr::slice_tail(the_tate_area)
```

The artist with the largest work in the tate collection is Therrien, Robert

The artist with the smallest work in the collection is Mesens, E.L.T.. The smallest work has area 2.37 $\text{cm}^2$

### Exercise 7

```{r}
joined_data <- the_tate %>% dplyr::left_join(the_tate_artists, by = join_by(artistId == id)) %>% tidyr::drop_na(gender)

joined_data <- joined_data  %>% dplyr::group_by(gender)

view(joined_data)
```

### Exercise 8

```{r}
spx_historical_data <- readr::read_delim("data/SPX_HistoricalData_1692322132002.csv", ",", escape_double = FALSE, trim_ws = TRUE)
view(spx_historical_data)
```

```{r}
#Create mutated tibble where daily return and variance are calculated fields and remove null
spx_historical_mutated <- spx_historical_data %>% dplyr::mutate(year = lubridate::year(mdy(Date))) %>% dplyr::arrange(desc(lubridate::date(mdy(Date)))) %>% dplyr::mutate(daily_return =log10(`Close/Last`/dplyr::lead(`Close/Last`))) %>% dplyr::mutate(variance = daily_return*daily_return) %>% tidyr::drop_na(daily_return)

#Take log sum as annual return and sqrt of sum of variance as annual deviation
spx_historical_summary <- spx_historical_mutated %>% dplyr::group_by(year) %>% dplyr::summarize(
  annual_return = sum(daily_return)*100,
  std_dev = sqrt(sum(variance))*100,
  .groups = "drop"
)
spx_historical_summary
```

```{r}
view(spx_historical_mutated)
```

The annual return in the SPX price in 2020 was -9.389577%.

The corresponding price volatility was 10.481097%.

### Exercise 9

```{r}
period_return <- sum(spx_historical_summary$annual_return)
period_volatility <- sqrt(sum(spx_historical_summary$std_dev^2))

new_row <- tibble::tibble(year = "Period",annual_return = period_return,std_dev = period_volatility)
spx_historical_summary <- spx_historical_summary %>% dplyr::mutate(year = as.character(year))
spx_historical_summary = bind_rows(spx_historical_summary,new_row)

spx_historical_summary <- gt(spx_historical_summary)
spx_historical_summary <- 
  spx_historical_summary %>%
  tab_header(
    title = "SPX Summary Table",
  )
spx_historical_summary
```


The period volatility was 21.184664%

### 
